### 2019-09-05 initial/notes.tex
090519

Now, a slightly more subtle question: is it possible for CS2 to still exist but
lose stability? Consider that for some $\epsilon\lesssim \epsilon_c$, the
equilibria lie at $\phi_{2, 4} = \pi/2 \pm \Delta \phi$, where $\epsilon_c -
\epsilon = \eta \sin I \frac{\Delta \phi^2}{2}$, then we consider perturbation
about these equilibria $\phi_{2, 4} = \pi/2 \pm \Delta \phi + \delta \phi$. The
linearized equations then read
\begin{subequations}
    \begin{align}
        \dot{\phi} &\approx -\delta \mu \mp \eta^2 \sin I \cos I \delta \phi,\\
        \dot{\mu} &\approx \pm\eta \sin I \Delta \phi \delta \phi.
    \end{align}
\end{subequations}
The eigenvalue equation for this first-order system reads
\begin{equation}
    \lambda^2 \pm \eta^2 \sin I \cos I \pm \eta \sin I \Delta \phi = 0.
\end{equation}
This has real solutions if
\begin{equation}
    \p*{\eta^2 \sin I \cos I}^2 \mp \eta \sin I \Delta \phi > 0.
\end{equation}
If $\epsilon \sim \epsilon_c / 2$, i.e.\ not too close to critical, then
$\Delta \phi^2 \sim 1$, so except for $\epsilon_c - \epsilon \ll \epsilon_c$,
the second term dominates the first and we have one real and one imaginary root,
corresponding to the two CSs. Otherwise, we can identify that for
\begin{align*}
    \Delta \phi &\leq \eta^3 \sin I \cos^2 I,\\
    \epsilon_c - \epsilon &\leq \frac{\eta^7 \sin^3 I \cos^4 I}{2},
\end{align*}
that CS2 becomes unstable. We can also numerically validate this via
\lstinline{99_misc/0_stab.py}.

@@

Consider if $\eta$ crosses $\eta_c$ with $\eta(t = 0) > \eta_c$, and there is no
dissipation, only variation of $\eta$. The area contained within the separatrix
when it first forms is given in Ward \& Hamilton Paper I\@:
\begin{equation}
    A_{crit} = 4\pi \s*{1 - \p*{1 + \tan^{2/3}I}^{-3/2}}.
\end{equation}
We are interested in states that start nearly aligned w/ the disk, so in Zone II
by our designation. Then, once three states appear, Zone II continues to shrink
until separatrix crossing occurs. At the time of separatrix crossing, the
trajectory is ejected into either Zone I/III (if Zone III is shrinking, then
exclusively Zone I\@; can analytically state using Henrard \& Murigande 1987).
The possible values for the new action of the ejected state are either $A_2 +
A_3 - 2\pi$ or $A_3 - 2\pi$, when integrating $\mu\;\mathrm{d}\phi$. From here,
the states will precess uniformly at some $\theta_f = A / 2\pi$.

>>>>

Now, a slightly more subtle question: is it possible for CS2 to still exist but
lose stability? Consider that for some $\epsilon\lesssim \epsilon_c$, the
equilibria lie at $\phi_{2, 4} = \pi/2 \pm \Delta \phi$, where $\epsilon_c -
\epsilon = \eta \sin I \frac{\Delta \phi^2}{2}$, then we consider perturbation
about these equilibria $\phi_{2, 4} = \pi/2 \pm \Delta \phi + \delta \phi$. The
linearized equations then read
\begin{subequations}
    \begin{align}
        \dot{\phi} &\approx -\delta \mu \mp \eta^2 \sin I \cos I \delta \phi,\\
        \dot{\mu} &\approx \pm\eta \sin I \Delta \phi \delta \phi.
    \end{align}
\end{subequations}
The eigenvalue equation for this first-order system reads
\begin{equation}
    \lambda^2 \pm \eta^2 \sin I \cos I \pm \eta \sin I \Delta \phi = 0.
\end{equation}
This has real solutions if
\begin{equation}
    \p*{\eta^2 \sin I \cos I}^2 \mp \eta \sin I \Delta \phi > 0.
\end{equation}
If $\epsilon \sim \epsilon_c / 2$, i.e.\ not too close to critical, then
$\Delta \phi^2 \sim 1$, so except for $\epsilon_c - \epsilon \ll \epsilon_c$,
the second term dominates the first and we have one real and one imaginary root,
corresponding to the two CSs. Otherwise, we can identify that for
\begin{align*}
    \Delta \phi &\leq \eta^3 \sin I \cos^2 I,\\
    \epsilon_c - \epsilon &\leq \frac{\eta^7 \sin^3 I \cos^4 I}{2},
\end{align*}
the eigenvalues become real. However, for CS2, since the eigenvalue equation
reads $\lambda^2 + b\lambda + c = 0$ and the eigenvalues take form $\p*{-b \pm
\sqrt{b^2 - 4c}}/2$, no eigenvalue w/ positive real part ever appears, so it
remains forever stable. We can also numerically validate this via
\lstinline{99_misc/0_stab.py}.

@@

Consider if $\eta$ crosses $\eta_c$ with $\eta(t = 0) > \eta_c$, and there is no
dissipation, only variation of $\eta$. The area contained within the separatrix
when it first forms is given in Ward \& Hamilton Paper I\@:
\begin{equation}
    A_{crit} = 4\pi \s*{1 - \p*{1 + \tan^{2/3}I}^{-3/2}}.
\end{equation}
We are interested in states that start nearly aligned w/ the disk, so in Zone II
by our designation. Then, once three states appear, Zone II continues to shrink
until separatrix crossing occurs, which is at $J_i = A_2$. At the time of
separatrix crossing, the trajectory is ejected into either Zone I/III (if Zone
III is shrinking, then exclusively Zone I\@; can analytically state using
Henrard \& Murigande 1987). The possible values for the new action of the
ejected state are either $J_f = A_2 + A_3 - 2\pi$ or $J_f = A_3 - 2\pi$, when
integrating $\mu\;\mathrm{d}\phi$. From here, the states will precess uniformly
at some $\theta_f = J_f / 2\pi$.

We can investigate two behaviors here then, the range of permitted $\theta_f$ as
a function of $I$ (when $J_i = A_{crit}$), and the values of $\theta_f$ given
some initial mutual inclination $\hat{s} \cdot \hat{l}_p$ ($J_i = 2\pi\p*{1 -
\cos \epsilon}$ where $\epsilon$ is the initial mutual inclination).
